### 2009-09-04 latex/paper1/contents.tex

Unfortunately, in the case of legacy systems, such knowledge isn't available.
Depending on its size, it might take months to understand the system so well as
to find a good modularization. % CITE Parnas' modules

@@

The empirical evaluations consist of selecting a collection of systems with
known reference modularizations and then applying the algorithms to the systems.
The modularizations found by the algorithms are then compared to the reference
decompositions by a metric such as MoJo CITE or PrecisionRecall CITE.

@@

It's an task of reverse engineering, as defined by Tonella...

@@

\section{Complex Networks}

Complex network theory found many scale-free networks

Software dependency networks are scale free. CITE Valverde, Myers

Scale free means ... $N(k) \sim k^{-\gamma}$

@@

Many scale-free models have been proposed. Only a few, though, produce modular
networks.

@@

picks a event that evolves
the network. Each event has an associated probability.

* Add a node with ongoing edge. With probability p, a new vertex is added to the network, together with an
edge from the new vertex to an existing vertex, chosen preferentially according
to .... Considering din = 0, this means that a vertex with indegree = 8 is
twice more likely to receive the edge than a vertex with indegree = 4. The
parameter din can be used to alleviate the handicap. (If din = 4, the first
vertex will only be 3/2 more likely to receive the edge). The new vertex is
put on the same cluster as the old vertex.

* Add a node with ingoing edge. With probability q, a new vertex is added to the network, together with an
edge from an existing vertex, chosen preferentially accordingly to dout +
outdegree, to the new vertex. The new vertex is put on the same cluster as the
old vertex. This case is similar to the previous case.

* Add an edge. With probability r, a new edge is added between two vertices, v and w. v is
chosen according to .... After that, with prob X, w is chosen from one of the
modules connected to v's module, else it's chosen from v's own module. In any
case, the exact choice is made according to ...

It's a growth model, that is, it can take as input an existing network and
evolve it.

TODO: Argument: why the imposed modules are natural modules?

@@

\section{Experimental Setup}

@@

``softwareness''

Our hypothesis is that at least one of the presented models can synthesize
networks that resemble software networks. We already know that they produce
networks that, just like software networks, are scale-free. This single
property, though, isn't enough to prove the hypothesis, since there are many
known scale-free networks extracted from many distinct domains of knowledge. It
is important to distinguish between software and non-software networks.

@@

We've collected 65 systems written in Java and X networks from many domains,
such as sociology, biology, technology and linguistics.

See Appendix A

@@

In a recent work, Milo et al. proposed to distinguish networks of different
domains by analyzing the triad concentration of each network. A triad is simply
a network with three vertices in which all vertices are connected. There are
only 13 distinct triads -- one for each configuration of directed edges --, as
shown in Figure X.
%CONSIDERING all the possible configurations of
%directed edges between three vertices, there are to 13 triads in which all
%vertices are connected, as shown in Figure X.

% Figure X: triad id vs. graphical representation

In order to characterize a network, one can count how many times each of the
13 triads appear on the network. The result is a triad concentration profile.
Figure 1a show triads for a software system... Figure 2a show triads for network
from domain X.

@@

In order two measure the softwareness of a network, we measure the correlation
between the network's profile and the profile of each of the 65 software
networks in our sample. We then compute the average of these correlations and
call it the S-score, that varies between 0 and 1. The greater the S-score, more
likely that the network is a software network.

@@

@@

\subsection{Network Synthesis}

In assessing the realism of a model, we want to know if, with a proper choice of
parameters, the model is capable of producing a realistic network. In order to
do so, we must generate many networks using different combinations of
parameters.

Since most of the parameters can assume infinite distinct values, we chose to
fix some of them and vary the others in discrete steps. In all models the number
of vertices was fixed to 1000. We generated one network for each set of
parameters. Here we describe the criteria we use to choose the parameters for
each model.

We chose five different module dependency networks, which where extracted from
actual dependencies between modules of five different software systems of our
sample, ranging from 2 to 32 modules. The module dependency networks are shown
on Figure XXX. 
%02-gef
%04-ibatis
%08-megamek
%16-findbugs
%32-zk

The probabilities p, q and r were given all possible values from 0.0 upto 1.0,
in 0.2 steps, such as the sum of the probabilities was 1. Since the only events
that create vertices are those associated with probabilities p and q, we imposed
the additional restriction that p + q > 0.

@@

Show some graphs: histogram of average correlations for each model.

@@

\section{Conclusion and Future Work}

We have shown empirically that network models found in the literature can
produce synthetic networks that are similar to the network of dependencies
between classes in object-oriented systems. This result supports the use of
synthetic networks in the evaluation of software modularization algorithms that
operate on class dependency networks.

% review
The use of synthetic data is common in distributed computing research, but still
underexplored in software engineering research. Since many reverse engineering
tasks extract from component dependencies information useful for the maintenance
of a software system, we expect this work to have impact beyond the software
modularization recovery community.

@@

We accept that it is very important to test the algorithms with networks
extracted from real software systems, but we argue that the use of synthetic
networks in a complementary manner can give researchers new insights about the
algorithms. First, the use of models allows the creation of large test sets,
thus diminishing the small sample effects. Moreover, the networks are created in
a controlled way, accordingly to model parameters, so it is possible to study
the behavior of the algorithms with different parameter values.

In a future work, we intend to evaluate with synthetic networks some software
clustering algorithms that were previously studied with real data. We will,
then, be able to compare the results obtained with the two approaches.

>>>>

Unfortunately, in the case of legacy systems, such knowledge isn't available.
Depending on its size, it might take months to understand the system so well as
to find a good modularization. XXX \cite{Parnas1972}

@@

The empirical evaluations consist of selecting a collection of systems with
known reference modularizations and then applying the algorithms to the systems.
The modularizations found by the algorithms are then compared to the reference
decompositions by a metric such as MoJo \cite{Tzerpos1999} or PrecisionRecall.

  \begin{figure*}[!t]
  \centering
  \includegraphics[width=1.0\textwidth]{diagram}
  \caption{Evaluation of a software modularization recovery algorithm}
  \label{fig:diagram}
  \end{figure*}

@@

It's an task of reverse engineering, as defined by Tonella \cite{Tonella2007}...

DEFINITIONS

external edge...

@@

\section{Network Theory}

Network theory research studies general properties of many types of networks by
using statistical analysis. In the last decade, it has been found that many
networks arising from sociology, biology, technology and other domains present
remarkable structural similarities. It has been shown that, in these networks,
the number of vertices connected to k edges, N(k), is proportional to
$k^{\-gamma}$, where $\gamma$ is a positive constant. These networks were called
scale-free networks.

Network theory has been applied to software networks and it was shown that they
are also scale-free networks \cite{Myers2003,Valverde2003}.

@@

Many models were proposed to explain the formation of scale-free networks. These
models are simple algorithms that can be proven, either formally or empirically,
to generate networks that are scale-free.

@@

Before we present the algorithm in detail, using pseudo-code, we must explain
some definitions. V is the set of vertices in the network being generated. This
set grows as the algorithm is executed. Being v a vertex, we define the
following three functions:

* out-neighbors(v): the set of all vertices that are connected to v by an edge
starting in v.
* same-module(v): the set of all vertices that are in the same module as v (except
v itself).
* other\_modules(v): the set of all vertices that are in modules that are connected
to v's module (in the graph G) by an edge starting in v's module.

After the initial network is created, the algorithm proceeds as follows:

while the network has less than n vertices:
  choose one of the operations according to probabilities (p, q, r)
  operation 1: (add a vertex with an outgoing edge)
    choose a vertex w within V according to f(x) = din + in-degree(x)
    add a new vertex v to w's module
    add an edge from v to w
  operation 2: (add a vertex with an ingoing edge)
    choose a vertex w within V according to f(x) = dout + out-degree(x)
    add a new vertex v to w's module
    add an edge from w to v
  operation 3: (add an edge between two existing vertices)
    choose a vertex v within V according to f(x) = dout + out-degree(x)
    choose a case according to probabilities (mu, 1 - mu)
    case 1: (distinct modules)
  	choose a vertex w within (other\_modules(v) - neighbors(v)) according to f(x) = din + in-degree(x)
      add an edge from v to w
    case 2: (same module)
      choose w within (same-module(v) - neighbors(v)) according to f(x) = din + in-degree(x)
      add an edge from v to w
  end

We can see that the probabilities p, q and r control how often each operation is
performed. Because the operation associated with probability r does not add any
vertices, greater values of r imply more edges in the resulting network. It is
easy to see that nodes connected to many ingoing edges are more likely to get
another ingoing edge, and the same reasoning can be applied to outgoing edges.
The parameter din can alleviate the handicap by providing a ``base in-degree''
that is applied to all vertices when computing the probabilities. Consider
two vertices, v1 with in-degree 4, and v2 with in-degree 8. If din = 0, v2 is
twice more likely to receive a new incoming edge; if, otherwise, din = 4, v2 is
only 3/2 more likely to receive the edge.

The parameter mu controls the proportion of edges between vertices in distinct
modules. Lower values of mu lend to networks that are more modular.

The BCR+ model is a growth model, meaning that the network is generated vertex
by vertex, growing from an initial network. It can, therefore, simulate the
evolution of a software network. Moreover, it can simulate the evolution of a
software system subject to constraints in module interaction, as is the case
with top-down design methodologies CITE.

%* Add a node with ongoing edge. With probability p, a new vertex is added to the network, together with an
%edge from the new vertex to an existing vertex, chosen preferentially according
%to .... Considering din = 0, this means that a vertex with indegree = 8 is
%twice more likely to receive the edge than a vertex with indegree = 4. The
%parameter din can be used to alleviate the handicap. (If din = 4, the first
%vertex will only be 3/2 more likely to receive the edge). The new vertex is
%put on the same cluster as the old vertex.
%
%* Add a node with ingoing edge. With probability q, a new vertex is added to the network, together with an
%edge from an existing vertex, chosen preferentially accordingly to dout +
%outdegree, to the new vertex. The new vertex is put on the same cluster as the
%old vertex. This case is similar to the previous case.
%
%* Add an edge. With probability r, a new edge is added between two vertices, v and w. v is
%chosen according to .... After that, with prob X, w is chosen from one of the
%modules connected to v's module, else it's chosen from v's own module. In any
%case, the exact choice is made according to ...
%
%It's a growth model, that is, it can take as input an existing network and
%evolve it.
%
%TODO: Argument: why the imposed modules are natural modules?

@@

The LF model is a very flexible model that can generated weighted directed
networks with overlapping modules, that is, in which a vertex can belong to more
than one module. Unlike the previous models, this is not a growth model: all
vertices are generated at once and then the edges are added.

There is also a special version of the model in which the edges weights are
discarded and the modules are non overlapping. We used the original
implementation of this version, available at http://.

\section{Experimental Setup}
% TODO: use simple past tense

@@

Our research hypothesis is that at least one of the presented models can
synthesize networks that resemble software networks. In order to investigate the
hypothesis we conducted an experiment. First we generated many synthetic
networks, using all three models and different combinations of parameters.
After that we selected the networks that were more similar to software networks.
Finally, we sought patterns in the parameters of the selected networks.

We assume that software networks possess a intrinsic quality, that we will call
``softwareness'', that distinguish them from networks in other domains. A
measure of the softwareness is essential to the experiment. One possible measure
is described and evaluated in the following sections. The measured softwareness
of a network represents how likely it is a software network.

We already know that they
produce networks that, just like software networks, are scale-free. This single
property, though, isn't enough to prove the hypothesis, since there are many
known scale-free networks extracted from many distinct domains of knowledge.

To support the analysis of softwareness

@@

We have collected 131 networks to support the experiment. The networks are
described in detail in Appendix A.

Software networks. We have collected 65 software systems written in Java. Java
is a popular programming language, for which there is plenty of open source
software.  The corresponding software networks were extracted by the tool
Dependency Finder CITE, considering classes and interfaces as components.
Dependency Finder was used because it can be easily integrated with scripts.

Non-software networks. We have collected 66 networks from distinct domains, such
as biology, sociology, technology, and linguistics. These networks are freely
available on the Internet and have been used in previous researches.

%We collected 65 software systems written in the Java
%programming language and 66 networks from other domains, such as biology,
%sociology and linguistics. Both the software systems and the networks are freely
%available on the Internet.
%
%As of this writing, Java is a popular programming language, for which there is
%plenty of open source software CITE. Also, there are many dependency extractors
%for this language. We have extracted the dependency network for each software
%system using a tool called Dependency Finder CITE, for its easy integration with
%scripts. Dependency Finder looks for dependencies by means of static analysis of
%Java bytecode and outputs the dependency network in a XML file. Although this
%network is in a detailed level, showing dependencies between fine-grained
%implementation entities such as attributes and methods, we have abstracted it to
%a network of dependencies between classes. The network is directed and
%non-weighted. The size of the networks range from XXX to YYY.

% JARs?

%The networks from other domains were collected in websites of complex network
%research groups. The domains include sociology, linguistics, biology and
%technology. The size of the networks range from XXX to YYY vertices.

\subsection{Synthetic Data Set}

We want to investigate if, with a proper choice of parameters, a model is
capable of synthesizing a network that resembles software networks. In this
section we describe which combinations of parameters values were used for each
model.

%In assessing the realism of a model, we want to know if, with a proper choice of
%parameters, the model is capable of producing a realistic network. In order to
%do so, we must generate many networks using different combinations of
%parameters.

In order to keep the networks in a tractable number, we have set the number of
vertices in each network to 1000, and varied the remaining parameters in
discrete steps such as each parameter assumed no more than 6 values. We have
generated one network for each combination of parameter values.

%Since most parameters are continuous, we had to 
%we chose to threat the number of vertices
%as a constant and vary the other parameters in discrete steps. In all models the
%number of vertices was fixed to 1000. We generated one network for each set of
%parameters.
%Here we describe the criteria we use to choose the parameters for each model.

We chose five different module dependency networks, which where extracted from
actual dependencies between archives of five different software systems of our
sample. The module dependency networks are shown on Figure
\ref{fig:architectures}.

\begin{figure*}[!t]
\centering
\includegraphics[width=1.0\textwidth]{architectures}
\caption{Module dependency graphs}
\label{fig:architectures}
\end{figure*}

%02-gef
%04-ibatis
%08-megamek
%16-findbugs
%32-zk

The probabilities p, q and r were given all possible values from 0.0 up to 1.0,
in 0.2 steps, such as the sum of the probabilities was 1. Since the only events
that create vertices are those associated with probabilities p and q, we imposed
the additional restriction that p + q > 0.

@@

In a recent work, Milo et al. proposed to distinguish networks of different
domains by analyzing the triad concentration of each network. A triad is simply
a network with three vertices in which all vertices are connected. There are
only 13 distinct triads -- one for each configuration of directed edges --, as
shown in Figure \ref{fig:triads}.
\cite{Milo2002}
%CONSIDERING all the possible configurations of
%directed edges between three vertices, there are to 13 triads in which all
%vertices are connected, as shown in Figure X.

\begin{figure*}[!t]
\centering
\includegraphics[width=1.0\textwidth]{triads}
\caption{Triads (id vs. graph. reprs.)}
\label{fig:triads}
\end{figure*}

In order to characterize a network, one can count how many times each of the
13 triads appear on the network. The result is a triad concentration profile.
Figure 1a show triads for a software system... Figure 2a show triads for network
from domain X. \ref{fig:profiles}

\begin{figure*}[!t]
\center
\subfigure[ref1][Software
network]{\includegraphics[width=2.5in]{triads-software}}
\hfil
\subfigure[ref2][Non-software network]{\includegraphics[width=2.5in]{triads-nonsw}%
}
\caption{Triad profiles}
\label{fig:profiles}
\end{figure*}

@@

CITE igraph

In order two measure the softwareness of a network, we measure the correlation
between the network's profile and the profile of each of the 65 software
networks in our sample. We then compute the average of these correlations and
call it the S-score, that varies between -1 and 1. The greater the S-score, more
likely that the network is a software network.

@@

TODO: show histogram of S-scores for sw networks (x axis from -1.0 to 1.0)
TODO: Is there need for cross validation in order to choose a threshold?
 - do cross-validation and choose threshold with max precision

@@

@@

Show some graphs: histogram of average correlations for each model.
\ref{fig:histograms}

\begin{figure*}[!t]
\center

\subfigure[Software network]{\includegraphics[width=1.0in]{hist-cgw}}

\subfigure[Non-software network]{\includegraphics[width=1.0in]{hist-bcr}}

\subfigure[Non-software network]{\includegraphics[width=1.0in]{hist-lf}}

\caption{Triad profiles}
\label{fig:histograms}
\end{figure*}

@@

\section{Conclusion and Future Work}
% revised 2009-09-03

We have shown empirically that network models found in the literature can
synthesize networks that resemble the network of dependencies between classes in
object-oriented systems. This result supports the use of synthetic networks in
the evaluation of software clustering algorithms.
%that operate on class dependency networks.

The use of synthetic data is common in distributed computing research, but still
underexplored in software engineering research. Because many reverse engineering
tasks rely on dependency data, we expect this work to have impact beyond the
software clustering community.

@@

We accept that it is important to evaluate the algorithms with real software
networks, but we argue that the use of synthetic networks in a complementary
manner can give researchers new insights about the algorithms. First, the use of
models allows the creation of large test sets, thus diminishing the small sample
effects. Moreover, the networks are created in a controlled way, according to
model parameters, so it is possible to study the behavior of the algorithms with
different parameter values.

In a future work, we intend to use synthetic networks in the evaluation of
software clustering algorithms that were previously studied with real networks
\cite{Wu2005}. After that we will be able to compare the results obtained by
both approaches.
